---
title: "Class 1"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisites

Install needed packages

```{r install_packages, echo=TRUE, exercise=FALSE}
pck <- c('devtools','stringr')
install.packages(pck[!pck %in% rownames(installed.packages())])
invisible(lapply(pck,library,character.only = TRUE, quietly = TRUE))
```

Install working RandomFieldsUtils and Miraculix versions

```{r install_dependencies, eval=FALSE, exercise=FALSE, include=TRUE}
pck <- character(2)
temp <- readLines('https://github.com/tpook92/MoBPS')
pck[1] <- str_extract(grep('RandomFieldsUtils_[0-9.]+tar.gz',temp, value = TRUE),
                      'RandomFieldsUtils_[0-9.]+tar.gz')[[1]]
pck[2] <- str_extract(grep('miraculix_[0-9.]+tar.gz', temp, value = TRUE),
                      'miraculix_[0-9.]+tar.gz')[[1]]
pck <- paste0('https://github.com/tpook92/MoBPS/raw/master/', pck)
install.packages(pck, repos = NULL, type = 'source')


```


Install MoBPS

```{r install_MoBPS, echo=TRUE, exercise=FALSE}
devtools::install_github('tpook92/MoBPS/pkg')
library(MoBPS)
```

## Creation of a base population

MoBPS is based on the assumption that every breeding program can be modelled by defining ....

We will start by simulating a population of 10 individuals and 100 SNPs. The basic function to do so is `MoBPS::creating.diploid()`, which creates the initial population list of founder animals. There also exists a tailored `summary()` function, printing the most necessary information. 

```{r show_base_creation, echo=TRUE}
popList <- MoBPS::creating.diploid(nindi = 10,
                                   nsnp = 100)
summary(popList)
```

As you noticed, MoBPS automatically divided the population into 50% male and 50% female individuals and assigned them to seperate cohorts. So the population consists of two cohorts.

If you want to specify which individuals you want to add, use the parameter `sex.s` instead of `nindi`. It needs a vector of `1` (male) and `2` (female). So following code creates the same population as before, just all individuals being male. 

```{r show_base_creation_2, echo=TRUE}
popList <- MoBPS::creating.diploid(sex.s = rep(1,10),
                                   nsnp = 100)
summary(popList)
```


### exercise 1
Now it's your turn. Create a base population `basePop` of 10 males, 50 females and 1000 SNPs.

```{r creation, exercise=TRUE, exercise.lines = 5}
# Create a base population basePop of 10 males, 50 females and 1000 SNPs

```


```{r creation-solution}
basePop <- MoBPS::creating.diploid(sex.s = c(rep(1,10), rep(2,50)),
                                   nsnp = 1000)
summary(basePop)
```


```{r eval=FALSE, include=FALSE}
<div id="creation-hint">
**Hint:** Use `rep()`to repeat values and combine them into one vector by `c()`.
</div>

```


```{r show_popList1, echo=TRUE}
str(popList, max.level = 1)
```

```{r show_popList2, echo=TRUE}
popList$info
```

```{r show_popList3, echo=TRUE}
popList$breeding
```



```{r}
popList <- breeding.diploid(popList)
str(popList$info)
```

