---
title: "Class 1 - basic functionality"
output: 
  learnr::tutorial:
    highlight: tango
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The following tutorial will lead you through the basic functionality of the R package MoBPS (Modular Breeding Program Simulator; <a href="https://www.biorxiv.org/content/10.1101/829333v1">www.biorxiv.org/content/10.1101/829333v1</a>) which can be found at <a href="https://github.com/tpook92/MoBPS">GitHub</a>.

If you do not know much about the intension and the general set up of the Simulation program yet, enjoy the presentation of Torsten in the following video.

![](https://youtu.be/uNeuCkWBNtg)


## Prerequisites

Install needed packages

```{r install_packages, echo=TRUE, exercise=FALSE}
pck <- c('devtools','stringr')
install.packages(pck[!pck %in% rownames(installed.packages())])
invisible(lapply(pck,library,character.only = TRUE, quietly = TRUE))
```

Install working RandomFieldsUtils and Miraculix versions

```{r install_dependencies, eval=FALSE, exercise=FALSE, include=TRUE}
pck <- character(2)
temp <- readLines('https://github.com/tpook92/MoBPS')
pck[1] <- str_extract(grep('RandomFieldsUtils_[0-9.]+tar.gz',temp, value = TRUE),
                      'RandomFieldsUtils_[0-9.]+tar.gz')[[1]]
pck[2] <- str_extract(grep('miraculix_[0-9.]+tar.gz', temp, value = TRUE),
                      'miraculix_[0-9.]+tar.gz')[[1]]
pck <- paste0('https://github.com/tpook92/MoBPS/raw/master/', pck)
install.packages(pck, repos = NULL, type = 'source')


```


Install MoBPS

```{r install_MoBPS, echo=TRUE, exercise=FALSE}
devtools::install_github('tpook92/MoBPS/pkg')
library(MoBPS)
```

## Creation of a base population

MoBPS is based on the assumption that every breeding program can be modelled by defining ....

The following video will introduce you to the content of this tutorial.

![](https://www.youtube.com/watch?v=xAl51woiU5s)

### Base population without phenotype

We will start by simulating a population of 10 individuals and 100 SNPs. The basic function to do so is `MoBPS::creating.diploid()`, which creates the initial population list of founder animals. There also exists a tailored `summary()` function, printing the most necessary information. 

```{r show_base_creation, echo=TRUE}
popList <- MoBPS::creating.diploid(nindi = 10,
                                   nsnp = 100)
summary(popList)
```

As you noticed, MoBPS automatically divided the population into 50% male and 50% female individuals and assigned them to seperate cohorts. So the population consists of two cohorts.

If you want to specify which individuals you want to add, use the parameter `sex.s` instead of `nindi`. It needs a vector of `1` (male) and `2` (female). So following code creates the same population as before, just all individuals being male. 

```{r show_base_creation_2, echo=TRUE}
popList <- MoBPS::creating.diploid(sex.s = rep(1,10),
                                   nsnp = 100)
summary(popList)
```


### exercise 1
Now it's your turn. Create a base population `basePop` of 10 males, 50 females and 1000 SNPs.

```{r creation, exercise=TRUE, exercise.lines = 5}
# Create a base population basePop of 10 males, 50 females and 1000 SNPs

```


```{r creation-solution}
basePop <- MoBPS::creating.diploid(sex.s = c(rep(1,10), rep(2,50)),
                                   nsnp = 1000)
summary(basePop)

basePop <- MoBPS::creating.diploid(nindi=60, sex.quota = 50/60,
                                   nsnp = 1000)
summary(basePop)
```


### Base Population with trait

```{r}
popList <- MoBPS::creating.diploid(nindi =  100,
                                   nsnp = 1000,
                                   trait.name = 'MKG',
                                   n.additive = 100)
summary(popList)
hist(popList$info$real.bv.add[[1]][,5],
     freq = FALSE, main = '', xlab = 'additive effect')
lines(density(popList$info$real.bv.add[[1]][,5]))
lines(seq(-3,3,0.01), dnorm(seq(-3,3,0.01)), col = 'red')
legend('topright',
       legend = c('simulated', 'standard normal'),
       col = c('black', 'red'),
       lty = 1)
```


```{r}
popList <- MoBPS::creating.diploid(nindi =  100,
                                   nsnp = 1000,
                                   trait.name = 'MKG',
                                   n.additive = 100,
                                   var.additive.l = 3) 
hist(popList$info$real.bv.add[[1]][,5],
     freq = FALSE, main = '', xlab = 'additive effect')
lines(density(popList$info$real.bv.add[[1]][,5]))
lines(seq(-5,5,0.01), dnorm(seq(-5,5,0.01)), col = 'red')
legend('topright',
       legend = c('simulated', 'standard normal'),
       col = c('black', 'red'),
       lty = 1)
```


```{r}
plot(popList$info$real.bv.add[[1]][,1],
     abs(popList$info$real.bv.add[[1]][,5]),
     main = '', xlab = 'position', ylab = 'absolute effect size')

```


### Adding second trait

```{r}
popList <- MoBPS::creating.diploid(nindi =  100,
                                   nsnp = 1000,
                                   trait.name = 'MKG',
                                   n.additive = 100,
                                   var.additive.l = 3) 

popList <- creating.trait(population = popList,
                          trait.name = c('FP'), 
                          n.additive = 200,
                          new.phenotype.correlation = matrix(c(1,-.5,-.5,1),2,2),
                          shuffle.traits = c(1,2),
                          shuffle.cor =  matrix(c(1,.5,.5,1),2,2))

# Correlation between QTL effects is obtained by placing effects on the same markers, leading to MORE underlying QTL than initially generated. This module is many intended for quantitative traits.
# In case a trait is simulated via no underlying QTL the genetic correlation can be entered via new.breeding.correlation
summary(popList)
str(popList$info,max.level = 2)
```

Be aware that using a genetic and phenotypic correlation matrix, which result in not defined environemental correlations will throw an error message. Note that such a matrix just can not occur in practice even those pair-wise correlation estimations might lead to such estimates:

```{r error=TRUE}
popList <- creating.trait(population = popList,
                          new.phenotype.correlation = matrix(c(1,-1,-1,1),2,2),
                          new.breeding.correlation = matrix(c(1,.5,.5,1),2,2))
```


```{r traits, exercise=TRUE, exercise.lines = 5}
popList <- MoBPS::creating.diploid(nindi =  100,
                                   nsnp = 1000)


```

```{r traits-solution}
# solution 1 --> create traits after population
popList <- MoBPS::creating.diploid(nindi =  100,
                                   nsnp = 1000)
popList <- MoBPS::creating.trait(population = popList,
                                 trait.name = c('traitA','traitB'),
                                 n.additive = c(150,20),
                                 var.additive.l = list(traitA = 1,traitB = 2),
                                 new.phenotype.correlation = matrix(c(1,0.5,0.5,1),2,2),
                                 shuffle.traits = 1:2,
                                 shuffle.cor = matrix(c(1,0.2,0.2,1),2,2))
summary(popList)
# solution 2 --> create traits directly
popList <- MoBPS::creating.diploid(nindi =  100,
                                   nsnp = 1000,
                                   trait.name = c('traitA','traitB'),
                                   n.additive = c(150,20),
                                   var.additive.l = list(traitA = 1,traitB = 2),
                                   new.phenotype.correlation = matrix(c(1,0.5,0.5,1),2,2),
                                   shuffle.traits = 1:2,
                                   shuffle.cor = matrix(c(1,0.2,0.2,1),2,2))
summary(popList)

```



## Breeding Actions


```{r}
popList <- MoBPS::creating.diploid(nindi =  100,
                                   nsnp = 1000,
                                   trait.name = 'MKG',
                                   n.additive = 100,
                                   var.additive.l = rep(3,100)) # change to single number when patched
popList <- MoBPS::breeding.diploid(popList,
                                   breeding.size = 100)
summary(popList)
```


```{r}
popList <- MoBPS::breeding.diploid(popList,
                                   breeding.size = 100,
                                   selection.m.gen = 2,
                                   selection.f.gen = 2,
                                   selection.criteria = 'bv',
                                   selection.size = c(10,50))
summary(popList)
```



## The population list object

```{r show_popList1, echo=TRUE}
str(popList, max.level = 1)
```

```{r show_popList2, echo=TRUE}
popList$info
```

```{r show_popList3, echo=TRUE}
popList$breeding
```



```{r}
popList <- breeding.diploid(popList)
str(popList$info)
```


## Evaluating results
